---
title: "Week 1 work"
author: "Arin Vansomphone"
date: "`r Sys.Date()`"
output: html_document
---
**Goal: To review on how to use survival models, learn JMBayes2 and how to use it**

## Review on survival models
Source: https://www.geeksforgeeks.org/survival-analysis-in-r/

Survival analysis is the prediction of events at a specified time. 

There are two methods of survival analysis in R:
* Kaplan-Meier 
* Cox proportional hazard model

### Kaplan-Meier 
Used for censored data, not based on underlying probability distribution
```{r}
library(survival)
lung
```

Fitting the model
```{r}
survival_function = survfit(Surv(lung$time, lung$status == 2) ~ 1)
survival_function
```

Plotting the function
```{r}
plot(survival_function, main = "Kaplan-Meier", xlab = "Number of days", 
     ylab = "Prob of survival")
```

### Cox proportional hazard model
Uses the hazard function, which considers independent variables in regression.

Also does not assume an underlying probability, but assumes that the hazards are constant over time.

More useful to measure instantaneous risk of deaths, often delivers better results because more volatile with data and features.

Tends to drop sharper as time increases. 
```{r}
cox_mod <- coxph(Surv(lung$time, lung$status == 2) ~., data = lung)

summary(cox_mod)
```

Fitting the model and plotting
```{r}
cox <- survfit(cox_mod)

plot(cox, main = "Cox proportional hazard model", xlab = "Number of days", 
     ylab = "Prob of survival")
```

## More advanced example of survival models
Source: https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/

Loading the data and packages
```{r}
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify) ## for better survival plots

data(veteran)
head(veteran)
```

### Kaplan Meier Analysis
```{r}
km <- with(veteran, Surv(time, status))
head(km, 80)
```

Fitting model with estimates at 1, 30, 60, and 90 days
```{r}
km_fit <- survfit(Surv(time, status) ~ 1, data = veteran)
summary(km_fit, times = c(1, 30, 60, 90*(1:10)))
```

Plotting
```{r}
autoplot(km_fit)
```

Survival curves by treatment
```{r}
km_trt_fit <- survfit(Surv(time, status) ~ trt, data = veteran)
autoplot(km_trt_fit)
```
 
Plotting patients by age
```{r}
vet <- mutate(veteran, AG = ifelse((age < 60), "LT60", "OV60"),
              AG = factor(AG),
              trt = factor(trt, labels = c("standard", "test")),
              prior = factor(prior, labels = c("NO", "Yes")))

km_AG_fit <- survfit(Surv(time, status) ~ AG, data = vet)
autoplot(km_AG_fit)
```

### Cox analysis
Fitting model
```{r}
cox <- coxph(Surv(time, status) ~ trt + celltype + karno + diagtime + age
             + prior, data = vet)
summary(cox)
```

Plotting
```{r}
cox_fit <- survfit(cox)
autoplot(cox_fit)
```

### Notes on cox regression models
* Need to be cautious when interpreting results because it assumes covariates are constant, and in many cases there are not

* Look more into Concordance stats if wanting to use ROC curves to assess model performance

Aalen's additive regression model shows how covariates change over time
```{r}
aa_fit <- aareg(Surv(time, status) ~ trt + celltype +
                  karno + diagtime + age + prior, data = vet)

aa_fit
```
```{r}
autoplot(aa_fit)
```

