---
title: "Week 3 Work"
author: "Arin Vansomphone"
date: "`r Sys.Date()`"
output: html_document
---

**Goal: Fit a longitudinal model onto the data and evaluate performance metrics**

```{r message=FALSE}
# load packages
library(readr)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(multilevelmod)
library(lme4)
library(jtools)
library(censored)
library(flexplot)

# load data
bloodpressure <- read_csv("bloodpressure.csv")
cvdoutcomes <- read_csv("cvdoutcomes.csv")
hba1c <- read_csv("hba1c.csv")
lipids <- read_csv("lipids.csv")
microvascularoutcomes <- read_csv("microvascularoutcomes.csv")
otherlabs <- read_csv("otherlabs.csv")
accordkey <- read_csv("accord_key.csv")
```

# Longitudinal Model of hba1c
```{r}
# removing unnecessary columns from accord key
accordkey <- accordkey %>%
  select(-c(trial_elig, network, arm))

# assembling covariates into one dataset
data <- inner_join(hba1c, lipids, 
                          by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
data <- inner_join(data, bloodpressure, 
                          by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
data <- inner_join(data, otherlabs, 
                          by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
data <- inner_join(data, accordkey, 
                          by = c('MaskID' = 'MaskID'))

data <- data %>%
  filter(Visit != "EXIT") # remove exit observations

# change visit to ordered factor
data[data == "BLR"] <- "00"
data$Visit <- gsub("F", "", data$Visit)
data$Visit <- as.numeric(data$Visit)
data <- data %>%
  mutate(Visit = factor(Visit, ordered = T))

# change some variables to factor
cols <- c("MaskID", "female", "treatment", "cvd_hx_baseline", "raceclass")
data[cols] <- lapply(data[cols], factor)
data
```

### Visualizing the data
```{r warning=FALSE}
data %>%
  filter(MaskID %in% c(100001, 100002, 100003, 100004, 100005, 100006)) %>%
  ggplot(aes(x = Visit, y = hba1c)) +
  geom_point() +
  geom_line() + 
  geom_path(group = 1) +
  facet_wrap(~MaskID)
```

```{r}
# fit model
full <- lmer(hba1c ~ Visit + chol + trig + vldl + ldl + hdl + sbp + dbp + hr + 
               fpg + alt + female + baseline_age + treatment + cvd_hx_baseline +
               raceclass + (1 | MaskID), # patient is random variable
           data = data)

require(flexplot)
estimates(full)
```

Note: figure out way to overcome missing data. Use baseline measurements for all observations? 

Note on design effect: the mixed model is equivalent to multiplying the dataset by $x$ factor (1 means mixed model does not contribute anything)

```{r}
# visualize model
flexplot::visualize(full, plot = "model",
                    formula = hba1c ~ Visit + MaskID,
                    sample = 100)
```

```{r}
# residual plot
flexplot::visualize(full, plot = "residuals")
```

# Creating a survival model
```{r}
# adding survival outcomes to dataset
# cvdoutcomes <- cvdoutcomes %>%
 # select(MaskID, fuyrs_po, censor_po)

#data <- inner_join(data, cvdoutcomes, 
                         # by = ('MaskID' = 'MaskID'))
```


```{r}
# splitting train and test
#data_train <- data[-c(1:5),]
#data_test <- data[1:5,]
```

```{r}
# creating survival model
# cox_spec <- proportional_hazards(penalty = 0.123) %>%
  # set_engine("glmnet")

# f_fit <- fit(cox_spec,
             # Surv(fuyrs_po, censor_po) ~ .,
             # data = data_train)

# f_fit
```

```{r}
# prediction
# f_pred <- predict(f_fit, new_data = data_test,
                 # type = "survival", time = seq(0, 20, 1))

# f_pred <- f_pred %>%
 # mutate(MaskID = factor(1:5)) %>%
 # unnest(cols = .pred)

#f_pred %>%
 # ggplot(aes(x = .eval_time, y = .pred_survival, col = MaskID)) +
  # geom_step()
```

# Week 3 summary

* Fit a baseline longitudinal model using `lme4`

* Determined methods of measuring performance through `flexplot`

**Next week's goals**

* Refine longitudinal model by adding parameters / looking into reduced model

* Build baseline survival model 
