---
title: "Week 5 Work"
author: "Arin Vansomphone"
date: "`r Sys.Date()`"
output: html_document
---
```{r message=FALSE}
# load packages
library(readr)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(multilevelmod)
library(lme4)
library(jtools)
library(censored)
library(flexplot)
library(survival)
library(ggfortify)
library(flexplot)
library(splines)
library(nlme)
library(performance)
library(effects)
library(lspline)
```

```{r message=FALSE}
# load data
bloodpressure <- read_csv("bloodpressure.csv")
cvdoutcomes <- read_csv("cvdoutcomes.csv")
hba1c <- read_csv("hba1c.csv")
lipids <- read_csv("lipids.csv")
microvascularoutcomes <- read_csv("microvascularoutcomes.csv")
otherlabs <- read_csv("otherlabs.csv")
accordkey <- read_csv("accord_key.csv")
```

# Improving hba1c longitudinal model 
```{r}
# removing unnecessary variables from accord key
accordkey <- accordkey %>%
  select(-c(trial_elig, network, treatment))

# assembling covariates into one dataset
df <- inner_join(hba1c, 
                 lipids, 
                 by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
df <- inner_join(df, 
                 bloodpressure, 
                 by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
df <- inner_join(df, 
                 otherlabs, 
                 by = c('MaskID' = 'MaskID', 'Visit' = 'Visit'))
df <- inner_join(df, 
                 accordkey, 
                 by = c('MaskID' = 'MaskID'))
df
```

```{r}
# remove exit observations
df <- df %>%
  filter(Visit != "EXIT")

# change visit to numeric
df[df == "BLR"] <- "00"
df$Visit <- str_replace(df$Visit, "F", " ")
df$Visit <- as.numeric(df$Visit)

# change some variables to factor
cols <- c("MaskID", "female", "cvd_hx_baseline", "raceclass", "arm")
df[cols] <- lapply(df[cols], factor)
df
```

```{r}
# count number of NAs in each column
sapply(df, function(x) sum(is.na(x)))
```

```{r}
# substitute covariates with many NAs to baseline only
baselinemeasurements <- df %>%
  filter(Visit == 0) %>%
  select(MaskID, cpk, potassium, screat, gfr, ualb, ucreat, uacr)

df <- inner_join(df, 
                 baselinemeasurements, 
                 by = ("MaskID"))

df <- df %>%
  select(-c(cpk.x, potassium.x, screat.x, ualb.x, gfr.x, ucreat.x, uacr.x)) %>%
  rename(bl_cpk = cpk.y, bl_potassium = potassium.y, bl_screat = screat.y, 
         bl_ualb = ualb.y, bl_gfr = gfr.y, bl_ucreat = ucreat.y, 
         bl_uacr = uacr.y)
df
```

```{r}
sapply(df, function(x) sum(is.na(x)))
```

Create small dataset for testing random effects and comparing models. 
```{r}
smalldf <- df %>% 
  filter(!MaskID %in% c("100026", "100028")) %>% # left out due to NAs
  head(196)
smalldf
```

## What variables should have a random effect?

### Arm
```{r}
# fit full model
noarmrandom <- lmer(hba1c ~ bs(Visit, degree = 1, knots = 12) + chol + trig + vldl + hdl + ldl + sbp + dbp + hr +
               fpg + alt + female + baseline_age + arm + cvd_hx_baseline +
               raceclass + bl_cpk + bl_potassium + bl_screat + bl_gfr + bl_ualb
             + bl_ucreat + bl_uacr + (1| MaskID), data = smalldf)
noarmrandom
```

```{r}
armrandom <- lmer(hba1c ~ bs(Visit, degree = 1, knots = 12) + chol + arm + trig + vldl + hdl + ldl + sbp + dbp + hr +
               fpg + alt + female + baseline_age + cvd_hx_baseline +
               raceclass + bl_cpk + bl_potassium + bl_screat + bl_gfr + bl_ualb
             + bl_ucreat + bl_uacr + (1 + arm | MaskID), data = smalldf)
armrandom
```

```{r}
# anova test
compare_performance(noarmrandom, armrandom, rank = T) 
```

Conclude that arm should be a fixed variable only. 

### Visit
```{r}
visitrandom <- lmer(hba1c ~bs(Visit, degree = 1, knots = 12) + chol + trig + vldl + hdl + ldl + sbp + dbp + hr +
               fpg + alt + female + baseline_age + arm + cvd_hx_baseline +
               raceclass + bl_cpk + bl_potassium + bl_screat + bl_gfr + bl_ualb
             + bl_ucreat + bl_uacr + (1 + Visit | MaskID), data = smalldf)
visitrandom
```

```{r}
novisitrandom <- lmer(hba1c ~bs(Visit, degree = 1, knots = 12) + chol + trig + vldl + hdl + ldl + sbp + dbp + hr +
               fpg + alt + female + baseline_age + arm + cvd_hx_baseline +
               raceclass + bl_cpk + bl_potassium + bl_screat + bl_gfr + bl_ualb
             + bl_ucreat + bl_uacr + (1 | MaskID), data = smalldf)
novisitrandom
```

```{r}
compare_performance(visitrandom, novisitrandom, rank = T) 
```

Conclude that visit should have a random slope. 

## Which covariates to choose?
```{r}
full <- lmer(hba1c ~ bs(Visit, degree = 1, knots = 12) + chol + sbp + dbp +
               fpg + female + baseline_age + arm + raceclass + bl_cpk + bl_screat + bl_gfr
             + bl_ucreat + (1 + Visit | MaskID), data = smalldf)
full
```

```{r}
partial <- lmer(hba1c ~ bs(Visit, degree = 1, knots = 12) + sbp + dbp +
               fpg + female + baseline_age + raceclass + arm + bl_cpk + bl_screat + bl_gfr
              + bl_ucreat + (1 + Visit | MaskID), data = smalldf)
partial
```

```{r}
compare_performance(full, partial, rank = T) 
```

### Variables kept:

* Visit 
* bl_gfr
* bl_screat
* bl_cpk
* baseline_age
* female (gender)
* fpg
* sbp
* dbp
* Visit
* arm
* raceclass

### Variables disgarded:

* bl_uacr
* bl_ualb
* cvd_hx_baseline
* hr
* hdl
* ldl
* vldl
* trig
* chol
* bl_potassium 
* alt

```{r}
original <- lmer(hba1c ~ Visit + chol + trig + vldl + hdl + ldl + sbp + dbp + hr +
               fpg + alt + female + baseline_age + arm + cvd_hx_baseline +
               raceclass + bl_cpk + bl_potassium + bl_screat + bl_gfr + bl_ualb
             + bl_ucreat + bl_uacr + (1 | MaskID), data = smalldf)
original
```

```{r}
final <- lmer(hba1c ~ lspline(Visit, knots = c(12)) + sbp + dbp +
               fpg + female + baseline_age + raceclass + arm + bl_cpk + bl_screat + bl_gfr
              + bl_ucreat + (1 + Visit | MaskID), data = smalldf)
final
```

```{r}
compare_performance(original, final, rank = T)
```

```{r}
plot(compare_performance(original, final))
```

Our final model is significantly improved. 

```{r}
model_performance(final)
```

## Performance plots 

```{r}
# flexplot residual plot
flexplot::visualize(final, plot = "residuals")
```

```{r}
# from the performance package
check_model(final)
```

## Predictions for longitudinal model
```{r}
smalldf %>%
  filter(!is.na(hba1c)) %>%
  mutate(pred = predict(final)) %>%
  filter(MaskID %in% c(100001, 100002, 100012)) %>%
  ggplot(aes(x = Visit)) +
  geom_line(aes(y = hba1c, group = MaskID, color = "Actual")) +
  geom_line(aes(y = pred, group = MaskID, color = "Prediction")) +
  labs(x = "Visit (months)") +
  facet_wrap(~MaskID) +
  scale_color_manual(values = c("Actual" = "black",
                                "Prediction" = "purple"))
```

Question: why is spline not working in model?

# Create survival model with tidymodels
```{r}
# adding survival outcomes to dataset
cvdoutcomes <- cvdoutcomes %>%
  select(MaskID, fuyrs_po, censor_po)
cvdoutcomes$MaskID <- as.factor(cvdoutcomes$MaskID)

df <- inner_join(df, 
                   cvdoutcomes, 
                   by = ('MaskID' = 'MaskID'))
df
```

```{r}
# setting up engine
library(censored)

decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("censored regression")
```

```{r}
# splitting training and testing data
library(survival)

train <- df[-c(1, 8, 16),]
test <- df[c(1, 8, 16),]
```

```{r}
train
```

```{r}
test
```

```{r}
# fitting model
cox_spec <- proportional_hazards(penalty = 0.123) %>%
  set_engine("glmnet")

f_fit <- fit(cox_spec, Surv(fuyrs_po, 1-censor_po) ~ hba1c + sbp + dbp +
               fpg + female + baseline_age + raceclass + bl_cpk + bl_screat + bl_gfr + bl_ucreat + arm, data = train)
f_fit
```

```{r}
# prediction and plotting
f_pred <- predict(f_fit, new_data = test, 
                  type = "survival", time = seq(0, 7, 0.5))

f_pred <- f_pred %>%
  mutate(MaskID = factor(1:3)) %>%
  unnest(cols = .pred)

f_pred %>%
  ggplot(aes(x = .eval_time, y = .pred_survival, col = MaskID)) + 
  geom_step()
```

Why is my model producing the same curve for different patients?